---
title: Lab 7 Revised
author: "Alyssa Walter"
editor: 
  markdown: 
    wrap: 72
---

For this lab, you will be joining and filtering related data sets to
solve a murder mystery!

# Part 1: GitHub Workflow

At the top of the document (in the YAML) there is an `author` line that
says `"Your name here!"`. Change this to be your name and save your file
either by clicking on the blue floppy disk or with a shortcut (command /
control + s).

Be sure to [commit the files to your
repo](https://happygitwithr.com/existing-github-first#stage-and-commit).

Let's get started!

# Part 2: Some Words of Advice

-   Set chunk options carefully.

-   Make sure you don't print out more output than you need.

-   Make sure you don't assign more objects than necessary---avoid
    "object junk" in your environment.

-   Make your code readable and nicely formatted.

-   Think through your desired result **before** writing any code.

# Part 3: Finding the Killer

Northwestern University's Knight Lab wanted to help sharpen users'
database skills, so they created a murder mystery. Can you solve this
crime in SQL City??

The relational data you will be working with contains tables with
different pieces of information pertinent to the crime - people, social
media check-ins, driver's licenses, crime scene reports, police
interviews, and more!

## Access the Data

This code chunk will read in **all** of the tables of data for you.
Don't modify or remove this! Take some time to look at each file type so
that

```{r}
#| message: false
#| warning: false
#| label: read-in-data

library(tidyverse)

# If purrr is not detected, install the package
if (!"purrr" %in% installed.packages()) install.packages("purrr")

source("https://raw.githubusercontent.com/jcanner/stat_210_2025_website/main/labs/instructions/lab-7-setup.R")
```

## Solve the Crime

### Crime Scene Report

Detective Wickham reaches out to you...

> A crime has taken place and I need your help! There was a murder in
> SQL City sometime on January 15, 2018. Could you retrieve the crime
> scene report from the police department’s database and follow the
> clues to find the person responsible for the murder?!

**Step 1: Find the police report in `crime_scene_report`. Then used the
information in the report to move on to the next data source based on
the information learned.**

```{r first_lead}
# Code for looking at the relevant crime scene report.

crime_scene_report %>%
  filter(date =="20180115", city == "SQL City", type == "murder") %>% #filter for the date of the crime, city, and crime type
  pull(description) #pull the description

```

**Next Steps: Follow the evidence to the person responsible for the
murder, building a report as you go.** There are accomplices, some
knowingly and some unknowingly, but there is only one mastermind.

Solve the murder mystery, showing **all of your work in this document**.
Your document and code must be well organized, easy to follow, and
reproducible.

-   Use headers and written descriptions to indicate what you are doing.
-   Use `dplyr` verbs and `join` functions rather than just looking
    through the tables manually. Functions from `stringr` and
    `lubridate` will also be useful.
-   Use good code formatting practices.
-   Comment your code.
-   Cite any external sources you use to solve the mystery.

>We need to find our witnesses so that we can search for their interview records. But all we have is their streets and a first name.

```{r finding_witnesses}

#searching for witness 1: Northwestern Dr
person %>%
  filter(address_street_name == "Northwestern Dr") %>%  #filter for the witness's address
  filter(address_number == max(address_number))  #find information for the last address


#searching for witness 2: Annabel on Franklin Ave
person %>%
  filter(address_street_name =="Franklin Ave", str_detect(name, "Annabel ")) %>%  #filter for the witness's address and name (from her first name)
  pull(name, id) #pull the resident's name and ID
```
>Next we are looking through the witness interviews, so that we can learn any information about the suspect.


```{r witness_interviews}

witness_ids <- c(14887, 16371) # save the two witness ids

person %>%
  filter(id %in% witness_ids) %>% #Filter for the ids
  left_join(interview, by = c("id" = "person_id")) %>% #join the interview data
  select(name, transcript) #pull the names and transcripts

```
>Since a witness believes the suspect is a Get Fit Now "gold" member, we need to identify anyone with the matching gym membership ids and date Annabel saw them.

```{r identifying_suspects}

# searching for member names and dates
get_fit_now_check_in %>%
  filter(str_detect(membership_id, "48Z"), check_in_date == "20180109") %>% #filter for partial id number and date
  left_join(get_fit_now_member, by = c("membership_id" = "id")) %>% #join the member data
  select(name, membership_status, check_in_date) #pull the names, membership status, and date


```

>It seems both men are now suspects because they were both at the gym the day Annabel saw the murderer. Lets follow up on the licence plate number we have from our witness Morty. 

```{r licence_plate_matching}

#To match a person to the plate, we first need to crosscheck the licence plate ids of our 2 suspects with our partial plate
person %>%
  filter(name %in% c("Joe Germuska", "Jeremy Bowers")) %>% #filter for suspect names
  left_join(drivers_license, by = c("license_id" = "id")) %>% #left join with license data
  filter(str_detect(plate_number, "H42W")) %>% #search for partial plate number
  select(name, plate_number) #pull the names and plate numbers


```
>We have found that Jeremy Bowers has the matching licence plate number so he is our top suspect, but lets see if he was interviewed.

```{r suspect_interviews}
# Jeremy Bowers
interview %>%
  filter(person_id =="67318") %>% #filter for the witness's id
  pull(transcript) #pull the interview transcript
```

>Lets search for records of a woman matching Jeremy's description.

```{r finding_murder_client}
drivers_license %>%
  filter(height %in% c("65", "66", "67"), hair_color =="red", car_make =="Tesla", car_model=="Model S") %>% #filter for the woman's height, hair, and car
  pull(id) #pull the interview transcript
```

>Now we have three suspects for the woman who hired our hit-man. Lets find their person ids so we can search for their records in the facebook check-ins data later. 

```{r}
person %>%
  filter(license_id %in% c("202298", "291182", "918773")) %>% #filter for suspect licence ids
  select(name, id, license_id) #pull their ids and names
```

>Now out of these 3 women, lets see who attended the concert 3 times like Jeremy recalled in his interview. I considered using a map function to do this, but decided just using summarise would be cleaner with our 3 ids.

```{r concert_attendants}

facebook_event_checkin %>%
  filter(str_detect(date, "201712"), #filter for December 2017
         event_name == "SQL Symphony Concert",  #look for the concert event check-ins
         person_id %in% c("78881", "90700", "99716")) %>% #filter for the ids of our female suspects
  group_by(person_id) %>% #group by person
  summarise(times_attended = n()) #count the number of times each suspect attended the concert

```
> Only Miranda Priestly attended the concert 3 times in December, so she must be the woman who hired our hit-man (who should also be charged).

::: callout-note
## And the final suspect is...

*Miranda Priestly*
:::
